### 2022-12-10 satellite-health-data-packet-as-table.tex
deploy: a4a77b6eaea83c04292b9bfae15b400f03ddbec3

\begin{tabular}{ | m{3cm} | m{1cm}| m{1cm} | m{4cm} | } 
  \hline
  \textbf{Field} & Type & Item Type & Description \\
  \hline
  callsign & string &  & Identifying call sign for the Sealion mission. \\
  \hline
  batteryhealth & float &  & Percent value indicating the remaining charge of the batteries. \\
  \hline
  temperaturebattery & float &  & The temperature of the battery. Units in Kelvin. \\
  \hline
  mode & integer &  & Integer value indicating current mission mode. 0 = Safe, 1 = mission mode 1, 2 = mission mode 2, 3 = mission mode 3. \\
  \hline
  tledata & TLE &  & TLE data from orbit propogator at time of beacon. \\
  \hline

\end{tabular}